# Module 4: Assessing Error & Sensitivity {#C05_1 -}

1) Calculate the mean absolute error in the simulation using the observed temperature data (`ObservedTempData.xlsx` from the *data* folder in *Kinneret97*) and the simulated temperature data (the simulated temperature data is from the surface layer in `lake.csv`).

<center>
\begin{equation}
MAE = \frac{\sum_{i}\left | T_{i}^{sim}-T_{i}^{obs} \right |}{n}
(\#eq:05-1)
\end{equation}
</center>

2) Calculate the sensitivity of the modelled temperature to changes in water clarity (the light extinction coefficient, $K_w$) and wind speed (`wind_factor`). These can be found in `glm3.nml`. Try increasing and decreasing the default parameter value by 0.2 and see how much the output changes.

<center>
\begin{equation}
SI = \frac{(Output_{new}-Output_{original})/Output_{original}}{(Parameter_{new}-Parameter_{original})/Parameter_{original}}
(\#eq:05-2)
\end{equation}
</center>

3) Assess how sensitive the phytoplankton biomass is to water clarity:

<center>
```{r table05-1, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(hablar)
library(dplyr)
options(kableExtra.html.bsTable = F, knitr.kable.NA = '')
module8_2 <- read.csv("tables/module8/table2.csv", check.names=FALSE)
kable(module8_2, format = "html", escape = F, align = "c") %>%
  kable_styling(full_width = F, position = "center") %>%
  column_spec(1, width_min = "12em") %>%
  column_spec(2, width_min = "12em") %>%
  column_spec(3, width_min = "12em") %>%
  column_spec(4, width_min = "12em") %>%
  scroll_box(width = "100%",
             fixed_thead = FALSE)
```
</center>
<br>

4) Assess how sensitive the phytoplankton biomass is to wind speed:

<center>
```{r table05-2, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(hablar)
library(dplyr)
options(kableExtra.html.bsTable = F, knitr.kable.NA = '')
module8_3 <- read.csv("tables/module8/table3.csv", check.names=FALSE)
kable(module8_3, format = "html", escape = F, align = "c") %>%
  kable_styling(full_width = F, position = "center") %>%
  column_spec(1, width_min = "12em") %>%
  column_spec(2, width_min = "12em") %>%
  column_spec(3, width_min = "12em") %>%
  column_spec(4, width_min = "12em") %>%
  scroll_box(width = "100%",
             fixed_thead = FALSE)
```
</center>

